\chapter{Conclusion and Future Direction}

\section{Conclusion}
This work explored multiple hybrid and deep learning-based models to improve stock return prediction. The initial Hybrid LSTM-RLS model showed consistent one-step lag, which persisted across the DNS and ARIMA-LSTM Residual Integration frameworks, despite minor RMSE improvements. The Multi-Feature LSTM Framework improved input representation but failed to resolve the lag issue.

GARCH integration over ARIMAX residuals proved ineffective, consistently predicting near-zero volatility. The Stochastic Volatility model followed general return patterns but exhibited lags and deviations. Transformer GAN and RAGIC architectures, implemented with minimal tuning, resulted in high prediction errors.

The Ensemble Learning Framework, combining ARIMAX, tree-based models, and a MetaNet, delivered the best performance by eliminating lag and closely tracking return trends, despite some directional prediction errors. Overall, ensemble learning proved the most effective strategy for real-time stock return forecasting among all models tested.

\section{Future Direction}
Future work will focus on refining the RAGIC and Transformer GAN architectures, as well as exploring alternative GAN-based frameworks aimed at achieving lag-free and low-loss stock return predictions. Additionally, incorporating more sophisticated feature engineering and advanced model tuning strategies could further enhance prediction accuracy and address the remaining challenges identified in this study.


\chapter{Publications}

\section*{Publications Based on This Dissertation}

\begin{itemize}
    \item Dr. Jigarkumar Shah, Dr. Bhubon Chandra Mech, Parth Sharma, \textit{“Hybrid Approaches for Improving Stock Price Forecasting Using Co-Integration of LSTM Network and RLS Algorithm”}, \textbf{28th Nirma International Conference on Management (NICOM 2025)}, held from \textbf{January 8 to January 10, 2025}. The presentation was part of the sub-theme \textbf{Finance and Accounting}, under the track titled “\textit{Financial Technologies and Digitalization: Financial Analytics and Machine Learning \& Deep Learning Applications in Finance.}”
\end{itemize}

\section*{Publications under Other Guides}

\begin{itemize}
    \item Parth Sharma, Rajesh Kumar Mohapatra, Dev Patel, Ritesh Vyas, Santosh Satapathy, \textit{“Automated Classification of EEG Patterns from BrainEEG Signals for Predicting Sleep Deficiency”}, \textbf{5th International Conference on Machine Learning, IoT and Big Data (ICMIB 2025)}, held from \textbf{April 4 to April 6, 2025}, Pandit Deendayal Energy University, Gandhinagar, Gujarat.
\end{itemize}

\section*{Target Journals for Future Publication}

Based on the research conducted in this dissertation, the following journals have been identified as suitable targets for future publication:

\section*{1. The Journal of Finance and Data Science}
\begin{itemize}
    \item \textbf{Impact Factor:} 3.9
    \item \textbf{Scopus Indexed:} Yes
\end{itemize}

\section*{2. Data Science in Finance and Economics}
\begin{itemize}
    \item \textbf{Impact Factor:} 1.3
    \item \textbf{Scopus Indexed:} No
\end{itemize}

These journals have been selected based on their relevance to the subject areas of finance, time series forecasting, and data science.